---
title: "Data Summary and Table 1"
subtitle: "SSI Sales Data"
author: "Stutti (Advisor: Dr. Seals)"
date: last-modified
format:
  revealjs
course: Capstone Projects in Data Science
bibliography: references.bib # file contains bibtex for references
#always_allow_html: true # this allows to get PDF with HTML features
self-contained: true
execute: 
  echo: false
  warning: false
  message: false
editor: 
  markdown: 
    wrap: 72
---
```{r setup, include=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(scales)
options(repos = c(CRAN = "https://cloud.r-project.org/"))
library(treemap)
```

## Table 1 - SSI Sales Data {.smaller .scrollable}

<style type="text/css">
.tg  {border-collapse:collapse;border-spacing:0;border-style: solid;border-width: 1px;border-color: black;}
.tg td{border-color:black;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:12px;
  overflow:hidden;padding:10px 5px;word-break:normal;}
.tg th{border-color:black;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:13px;
  font-weight:normal;overflow:hidden;padding:10px 5px;word-break:normal;}
.tg .tg-bobw{font-weight:bold;text-align:center;vertical-align:bottom}
.tg .tg-7zrl{text-align:left;vertical-align:bottom}
.tg .tg-j6zm{font-weight:bold;text-align:left;vertical-align:bottom}
.tg .tg-8d8j{text-align:center;vertical-align:bottom}
</style>
<table class="tg"><thead>
  <tr>
    <th class="tg-7zrl"></th>
    <th class="tg-j6zm">Total Orders</th>
    <th class="tg-j6zm">Closed Short</th>
    <th class="tg-j6zm">Fulfilled</th>
  </tr></thead>
<tbody>
  <tr>
    <td class="tg-7zrl"></td>
    <td class="tg-j6zm">(n=7585)</td>
    <td class="tg-j6zm">(n=733)</td>
    <td class="tg-j6zm">(n=6852)</td>
  </tr>
  <tr>
    <td class="tg-j6zm">Top Customers</td>
    <td class="tg-7zrl"> </td>
    <td class="tg-7zrl"> </td>
    <td class="tg-7zrl"> </td>
  </tr>
  <tr>
    <td class="tg-7zrl">Smoothie Island</td>
    <td class="tg-8d8j">1701 (22.43%)</td>
    <td class="tg-8d8j">455 (62.07%)</td>
    <td class="tg-8d8j">1246 (18.18%)</td>
  </tr>
  <tr>
    <td class="tg-7zrl">Philly Bite</td>
    <td class="tg-8d8j">1556 (20.51%)</td>
    <td class="tg-8d8j">267 (36.43%)</td>
    <td class="tg-8d8j">1289 (18.81%)</td>
  </tr>
  <tr>
    <td class="tg-7zrl">PlatePioneers</td>
    <td class="tg-8d8j">1396 (18.40%)</td>
    <td class="tg-8d8j">143 (19.51%)</td>
    <td class="tg-8d8j">1253 (18.29%)</td>
  </tr>
  <tr>
    <td class="tg-7zrl">Berl Company</td>
    <td class="tg-8d8j">906 (11.94%)</td>
    <td class="tg-8d8j">5 (0.68%)</td>
    <td class="tg-8d8j">901 (13.15%)</td>
  </tr>
  <tr>
    <td class="tg-7zrl">DineLink Intl</td>
    <td class="tg-8d8j">589 (7.77%)</td>
    <td class="tg-8d8j">42 (5.73%)</td>
    <td class="tg-8d8j">547 (7.98%)</td>
  </tr>
  <tr>
    <td class="tg-j6zm">Top Products</td>
    <td class="tg-8d8j"></td>
    <td class="tg-8d8j"></td>
    <td class="tg-8d8j"></td>
  </tr>
  <tr>
    <td class="tg-7zrl">DC-01</td>
    <td class="tg-8d8j">1135 (14.96%)</td>
    <td class="tg-8d8j">345 (47.07%)</td>
    <td class="tg-8d8j">790 (11.53%)</td>
  </tr>
  <tr>
    <td class="tg-7zrl">TSC-PQB-01</td>
    <td class="tg-8d8j">1087 (14.33%)</td>
    <td class="tg-8d8j">389 (53.07%)</td>
    <td class="tg-8d8j">698 (10.19%)</td>
  </tr>
  <tr>
    <td class="tg-7zrl">TSC-PW14X16-01</td>
    <td class="tg-8d8j">848 (11.18%)</td>
    <td class="tg-8d8j">283 (38.61%)</td>
    <td class="tg-8d8j">565 (8.25%)</td>
  </tr>
  <tr>
    <td class="tg-7zrl">CMI-PCK-01</td>
    <td class="tg-8d8j">802 (10.57%)</td>
    <td class="tg-8d8j">288 (39.29%)</td>
    <td class="tg-8d8j">514 (7.50%)</td>
  </tr>
  <tr>
    <td class="tg-7zrl">PC-05-B1</td>
    <td class="tg-8d8j">745 (9.82%)</td>
    <td class="tg-8d8j">220 (30.01%)</td>
    <td class="tg-8d8j">525 (7.66%)</td>
  </tr>
  <tr>
    <td class="tg-j6zm">Top Distributors</td>
    <td class="tg-8d8j"></td>
    <td class="tg-8d8j"></td>
    <td class="tg-8d8j"></td>
  </tr>
  <tr>
    <td class="tg-7zrl">Ed Don &amp; Company - Miramar</td>
    <td class="tg-8d8j">210 (2.77%)</td>
    <td class="tg-8d8j">0 (0.00%)</td>
    <td class="tg-8d8j">210 (3.06%)</td>
  </tr>
  <tr>
    <td class="tg-7zrl">PFG- Gainesville</td>
    <td class="tg-8d8j">197 (2.60%)</td>
    <td class="tg-8d8j">0 (0.00%)</td>
    <td class="tg-8d8j">197 (2.88%)</td>
  </tr>
  <tr>
    <td class="tg-7zrl">Ed Don &amp; Company -&nbsp;&nbsp;&nbsp;Woodridge</td>
    <td class="tg-8d8j">186 (2.45%)</td>
    <td class="tg-8d8j">0 (0.00%)</td>
    <td class="tg-8d8j">186 (2.71%)</td>
  </tr>
  <tr>
    <td class="tg-7zrl">Ed Don &amp; Company - Mira&nbsp;&nbsp;&nbsp;Loma</td>
    <td class="tg-8d8j">180 (2.37%)</td>
    <td class="tg-8d8j">0 (0.00%)</td>
    <td class="tg-8d8j">180 (2.63%)</td>
  </tr>
  <tr>
    <td class="tg-7zrl">.Ed Don - Miramar</td>
    <td class="tg-8d8j">162 (2.14%)</td>
    <td class="tg-8d8j">0 (0.00%)</td>
    <td class="tg-8d8j">162 (2.36%)</td>
  </tr>
  <tr>
    <td class="tg-j6zm">Top Substrates</td>
    <td class="tg-bobw">Paper</td>
    <td class="tg-bobw">Plastic</td>
    <td class="tg-bobw">Bagasse</td>
  </tr>
  <tr>
    <td class="tg-7zrl">Revenue($103,826,286)</td>
    <td class="tg-8d8j">$54,838,585 (52.82%)</td>
    <td class="tg-8d8j">$40,336,669 (38.85%)</td>
    <td class="tg-8d8j">$4,350,337 (4.19%)</td>
  </tr>
  <tr>
    <td class="tg-j6zm">Quantity Ordered</td>
    <td class="tg-bobw">Min</td>
    <td class="tg-bobw">Mean</td>
    <td class="tg-bobw">Max</td>
  </tr>
  <tr>
    <td class="tg-7zrl">Total Ordered(1,971,237)</td>
    <td class="tg-8d8j">1</td>
    <td class="tg-8d8j">61.47</td>
    <td class="tg-8d8j">23,160</td>
  </tr>
  <tr>
    <td class="tg-j6zm">Unit Price</td>
    <td class="tg-bobw">Min</td>
    <td class="tg-bobw">Mean</td>
    <td class="tg-bobw">Max</td>
  </tr>
  <tr>
    <td class="tg-7zrl">Key Stats</td>
    <td class="tg-8d8j">$0.16 </td>
    <td class="tg-8d8j">$62.60 </td>
    <td class="tg-8d8j">$864.00</td>
  </tr>
  <tr>
    <td class="tg-j6zm">Total Price</td>
    <td class="tg-bobw">Min</td>
    <td class="tg-bobw">Mean</td>
    <td class="tg-bobw">Max</td>
  </tr>
  <tr>
    <td class="tg-7zrl">Key Stats</td>
    <td class="tg-8d8j">$4.92 </td>
    <td class="tg-8d8j">$3,430.74 </td>
    <td class="tg-8d8j">$143,084.74</td>
  </tr>
</tbody></table>

# Data Visualization1
```{r}
# Load necessary libraries
library(tidyverse)
library(dplyr)
library(ggplot2)

#Load data
SaleData <- read_csv('/Users/ss/Documents/HR Misc/Masters/IDC6940 Capstone in Data Science/Data/DataSet1.csv')

# Filter out returns/credits (Price <= 0)
SaleData1 <- SaleData %>%
  filter(UnitPrice > 0, TotalPrice > 0)

# List of products to exclude
exclude_products <- c("OUTBOUND SHIPPING", "Outbound Shipping", "Shipping Charge", "SHIPPING CHARGE")  

# 1: Bar chart: Top 20 Products by Qty Ordered
product_qty <- SaleData1 %>%
  group_by(Product) %>%
  summarise(TotalQuantityOrdered = sum(qtyOrdered, na.rm = TRUE)) %>%
  arrange(desc(TotalQuantityOrdered)) %>%
  slice(1:20)

ggplot(product_qty, aes(x = reorder(Product, TotalQuantityOrdered), y = TotalQuantityOrdered)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  theme_minimal() +
  coord_flip() +
  labs(title = "Top 20 Products by Total Quantity Ordered", x = "Product", y = "Total Quantity Ordered")
```
# Data Visualization2

```{r}
library(scales)
# 2: Create a pie chart for TotalSales by Substrate
substrate_sales <- SaleData1 %>%
  filter(!Substrate %in% NA) %>%
  group_by(Substrate) %>%
  summarise(TotalSubstrateSales = sum(TotalPrice, na.rm = TRUE)) %>%
  arrange(desc(TotalSubstrateSales)) 

# configure sales to currency
substrate_sales$legend_label <- paste(substrate_sales$Substrate, "-", dollar(substrate_sales$TotalSubstrateSales))

ggplot(substrate_sales, aes(x = "", y = TotalSubstrateSales, fill = legend_label)) +
  geom_bar(stat = "identity", width = 1) +  # Create bar chart
  coord_polar("y") +  # Convert bar chart to pie chart
  labs(title = "Proportion of Sales by Substrate", fill = "Substrate and Sales") +  # Customize title and legend
  theme_void()  # Remove axis labels and background
```

# Data Visualization3
```{r}
# 3: Create a treemap of total sales by product
# Set CRAN mirror
#options(repos = c(CRAN = "https://cloud.r-project.org/"))
# Install the package
#install.packages("treemap")
library(treemap)

product_sales <- SaleData1 %>%
  filter(!Product %in% exclude_products) %>%
  group_by(Product) %>%
  summarise(pSales = sum(TotalPrice, na.rm = TRUE)) %>%
  arrange(desc(pSales)) %>%
  slice(1:30)
# Create a treemap of total sales by product
treemap(product_sales,
        index = "Product",
        vSize = "pSales",
        title = "Treemap of Total Sales by Product",
        palette = "Pastel1") # "PuBu"
```
# Data Visualization4
```{r}
# 4: Create a histogram of totalprice across sales orders
sales_filter <- SaleData1 %>%
  filter(!Product %in% exclude_products) 
  # filter(TotalPrice > 2000) %>% # Remove values causing skew
  # filter(TotalPrice < 30000) # Remove values causing skew
  
Q1 <- quantile(sales_filter$TotalPrice, 0.25, na.rm = TRUE)
Q3 <- quantile(sales_filter$TotalPrice, 0.75, na.rm = TRUE)
IQR_value <- IQR(sales_filter$TotalPrice, na.rm = TRUE)

lower_bound <- Q1 - 1.5 * IQR_value
upper_bound <- Q3 + 1.5 * IQR_value

filtered_data_IQR <- sales_filter %>%
  filter(TotalPrice >= lower_bound & TotalPrice <= upper_bound)

ggplot(filtered_data_IQR, aes(x = TotalPrice)) +
  geom_histogram(binwidth = 100, fill = "lightblue", color = "black") +
  theme_minimal() +
  labs(title = "Distribution of Total Prices (Filtered by IQR)", x = "Total Price", y = "Frequency")
```
# Data Visualization5
```{r}
# 5: Create heatmap - Top distributors by Orders
top_dist <- SaleData1 %>%
  group_by(OPCO) %>%
  summarise(Order_Count = n_distinct(SalesOrderID)) %>%
  arrange(desc(Order_Count)) %>%
  slice(1:20)

ggplot(top_dist, aes(x = "", y = OPCO, fill = Order_Count)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "darkblue") +
  labs(title = "Heatmap of Distributors by Orders", y = "Distributors") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

# Data Visualization6
```{r}
# 7: Sales Over Time
library(lubridate)
#install.packages("scales")
library(scales)

# Convert DateFulfilled in date format
### SaleData1$DateFulfilled <- ymd(SaleData1$DateFulfilled) ### Did not work
SaleData1$parsed_dates <- parse_date_time(SaleData1$DateFulfilled, orders = c("ymd", "mdy", "dmy", "Ymd", "Y-m-d"))
SaleData1$parsed_dates <- ymd(SaleData1$parsed_dates)

# Identify which entries failed to parse
# failed_dates <- SaleData1[is.na(SaleData1$parsed_dates), "DateFulfilled"]
# unique(failed_dates)


# Summarize total sales by quarter
sales_over_time <- SaleData1 %>%
  mutate(nDate = floor_date(parsed_dates, "quarter")) %>%
  group_by(nDate) %>%
  summarize(TotalSales = sum(TotalPrice, na.rm = TRUE))

# Create a new variable for quarters
sales_over_time$Quarter <- quarter(sales_over_time$nDate)

# sum(is.na(SaleData1$parsed_dates))  # Count NA values

# Plot total sales over time with quarterly labels
ggplot(sales_over_time, aes(x = nDate, y = TotalSales)) +
  geom_line(color = "blue") +
  geom_point() +
  scale_x_date(date_labels = "Q%q %Y", date_breaks = "3 months") +
  # scale_x_date(date_labels = "Q%q %Y", date_breaks = "1 quarter") +
  # scale_x_date(labels = label_date("%Y Q%q"), date_breaks = "1 quarter") +  # Correctly format quarters
  labs(title = "Total Sales Over Time (Quarterly)", x = "Quarter", y = "Total Sales") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
